Rare-event sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body

### 2019-05-27 draft_p2.tex
Relecture generale #11 #7 #2 #13 #6

\begin{abstract}
\TL{  This study evaluates the relevance of rare-event sampling type approaches to make accessible the numerical simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
  The main idea is here to replace a long simulation by a set of much shorter simulations, running in parallel, wich corresponding dynamics are replicated or pruned in order to sample large-amplitude events more frequently. 
%
  Such techniques have been shown to be efficient for a wide range of problems in statistical physics, computer science, biochemistry, enabling the simulation of events far from reach from brute force sampling approaches.
  Application to fluid-structure interaction problems remains mainly open. 
%
  The drag experienced by a squared obstacle placed in a turbulent flow (in two dimensions) is taken as a representative case study to investigate the performance of two major sampling algorithms, namely the \ac{ams} and the \ac{gktl} algorithms.
  Practical evidence is given that the fast sweeping time of fluid structures has a drastic influence on the efficiency of the sampling with both algorithms.
  While is is shown that the \ac{ams} resampling does not lead to rare trajectories with siginificant run-time savings, we illustrate that the \ac{gktl} algorithm allows for the simulation of extreme fluctuations of the time-averaged drag, as well as accurate computation of the associated statistics, for instance return times.}
\end{abstract}

@@

% general comments on physical problem %
%
Turbulent flows are important in a variety of natural phenomena, industrial and civil applications with the characteristic feature to spontaneously develop intense and sporadic motions associated with extreme internal forces \cite{lesieur_book,donzis_sreenivasan_2010,Yeung}. ``Extreme'' refers here to fluctuations that can deviate from mean value by ${\cal{O}}(10)$ standard deviations. In engineering, the nature of such extreme dynamical events and their statistics are of crucial interest to predict excessive mechanical efforts. Such anomalous efforts can threaten the structural integrity of embedded structures.
%
From the viewpoint of (chaotic) dynamical systems, turbulence in fluids has to do with non-linearity and strong departure from statistical equilibrium \cite{KRAICHNAN}. The use of perturbative methods in identifying resonant interactions (among degrees of freedom) possibly responsible for extreme fluctuations is not successful. Therefore, simulation offers a practical means to gain physical insight into these events, quantify their intensity and estimate their frequency of occurrence. Nevertheless, this requires very long simulation since these events are also very rare. It is here our motivation to evaluate the practicality and efficiency of rare-event sampling techniques to alleviate this prohibitive computational cost. 
%

% general def on rare-event sampling %
%
Rare-event sampling is an umbrella term for numerical methods intended to sample preferentially regions of the phase space which would otherwise be rarely visited through brute-force simulation.
%

%
% next two paragraphs may be too technical -- Find a way to summarize them in a few sentences accessible to a broad audience in fluid mechanics? Notion of action will be vague for many readers
%
% this paragraph would certainly need to be reshaped!
%
In fluid turbulence, rare-event sampling has been approached mainly from simplified dynamics such as the one-dimensional Burgers' equation with a stochastic forcing \cite{bec_burgers_2007}. In that case, dynamics may be accounted for through an action that can be identified as analogous to a Gibbs measure for a system at equilibrium. This latter can be sampled by using a Markov chain Monte-Carlo  algorithm \cite{duben_monte_2008,mesterhazy2011anomalous,mesterhazy2013lattice} that provides a framework for rare-event sampling.
%
An alternative approach is based on \emph{instantons} \cite{grafke2015instanton}. For stochastically driven systems in the limit of weak noise, it can be proved that the probability of extreme events is dominated by the minima of the action \cite{freidlin2013random} -- the trajectories that achieve such minima being referred to as instantons \cite{gurarie_instantons_1996}. Therefore, suitable numerical schemes can  be used to minimize the action and evaluate instantons as well as the related probabilities of rare events \cite{chernykh_large_2001,grafke_instanton_2013,grafke2015relevance,grigorio_instantons_2017}.
%
An obvious drawback of the aforementioned approaches is their limitation to stochastically driven dynamics which, in addition, must be simple enough so that an action may be derived and sampled, or minimized. 
%
%
%Whilst Burgers' turbulence provides an interesting model framework for the investigation of the properties of small scale turbulence, it is of little relevance to more realistic situations, such as industrial or environmental flows.
%
In this paper, a more general approach is investigated. It relies on sampling algorithms connected to the simulation of the actual dynamics itself, \textit{i.e.} applying directly to the trajectories in the dynamic space.
% dedicated to dynamical systems for which an action cannot be easily derived or minimized,
%    
%
Even though such ideas trace back to the early 1950's, they have received ever-growing interest over the last 20 years with successful applications in various domains such as chemistry \cite{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \cite{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \cite{louvin2017}, nonlinear dynamical systems \cite{tailleur_probing_2007} or communication networks simulation \cite{villen-altamirano_restart:_1994}.
%
% already mentioned in the beginning? This sentence is misleading.
%
%However, a large body of algorithms are limited to systems at equilibrium, for which the stationary measure can be sampled by means of Monte Carlo methods. 
%
An original contribution of the present study is certainly the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom. 
Let us remark that most of previous successful applications involve dynamical systems at statistical equilibrium and low-dimensional attractors. \EL{is it TRUE? \#11}
% 
Two different algorithms suitable for out-of-equilibrium dynamics are considered and compared. Namely, the Adaptive Multilevel Splitting algorithm and the Giardina-Kurchan-Tailleur-Lecomte algorithm.
%

@@

Over the last decade, the main theoretical framework for the study of rare events (in statistical physics) has been the theory of large deviations \cite{touchette_large_2009}.
Alongside, numerical methods have been developed to sample rare events \cite{moral_feynman-kac_2004}. 
Among them, the \ac{gktl} algorithm \cite{giardina_direct_2006} is particularly suited for (chaotic) dynamical systems \cite{giardina_simulating_2011,Laffargue_2013}.
%
Recently, \ac{gktl} has allowed to successfully perform the numerical simulation of extreme heat waves in a simplified modeling of the atmosphere \cite{ragone_computation_2018}. 
This achievement represents a significant leap in the applicability of rare-event sampling to complex dynamical systems. Along the same line, rare-event sampling algorithms are here pushed aside traditional applications by considering fluid-structure interaction in a turbulent flow.

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]{illustr_ecoulement/illustr_ecoulement}
\caption{Our case study is a grid-generated turbulent flow impinging onto a fixed squared obstacle (of size $R$) located at the center of a channel in two dimensions. The flow is artificially damped near the end of the channel. In the developed flow, turbulent eddies have typically the size of the square, which results in strong fluctuations of mechanical efforts acting on the square. The vorticity is displayed with an arbitrary color map from blue (negative values) to red (positive values).}
\label{fig:illustr_ecoulement}
\end{figure}

% introduce the flow 
%
The drag exerted by a grid-generated turbulent flow onto a fixed squared obstacle is considered as a representative case study (see Fig.~\ref{fig:illustr_ecoulement}). 
% why this flow?
%
Although real-world applications would eventually imply three-dimensional dynamics, a simplified two-dimensional setting has been chosen here to reduce the computational cost and allow for a systematic study. 
% 
We believe that this system embeds the characteristic features that makes the application of rare-event algorithms both relevant and challenging for fluid-structure-interaction problems. 
%Namely, spatio-temporal chaos and the emergence of large-amplitude mechanical efforts on the structure.
%
% rapid description of the flow
Turbulent eddies generated in the near-wake of the grid are carried downstream. They interact with each other and grow in size as expected in two-dimensional turbulence. 
The dimension of the grid is such that the size of the eddies that eventually hit the square is comparable to the size of this latter, which results in strong fluctuations of the drag acting on the square. 
%
%Finally, the obstacle does not deform or move either. 
%
%Through this simplified setting, our motivation is primarily to evaluate the operability of sampling techniques to capture extreme events with a significant run-time savings.  
%

@@

% geometry
%
The simulated flow develops in a long plane channel of dimension $512 \times 128$ mesh points. The squared obstacle has size $R=16$ (in mesh unit) and is located at the center of the channel. The spacing and bar height of the entrance grid are both equal to $R/2$ (see Fig.~\ref{fig:illustr_ecoulement}). 
% boundary conditions
%
No-slip boundary conditions are enforced on top and bottom walls of the channel and on the surface of the obstacle by using an halfway bounce-back procedure \cite{lbm_book}.  
%
Upstream the grid, a constant parabolic velocity profile and a constant mass density (equal to unity) are imposed as inlet condition. 
The centerline velocity is $0.05$ in lattice unit, \textit{i.e.} normalized by $\Delta x/\Delta t$ with $\Delta x$ and $\Delta t$ referring to the lattice resolution and the time-step respectively. The distributions are imposed at equilibrium (see Appendix). 
In the bulk, the viscosity is adjusted to generate grid turbulence with Reynolds number $\mathrm{Re_{grid}}=1200$. The reference Mach number is equal to $0.06$ in agreement with the assumption of weak compressibility of the LB method. 
Near the end of the channel, the flow is progressively damped within a \textit{sponge layer} where the viscosity is artificially enhanced. Finally, the oulet boundary condition relies on a second-order extrapolation of the velocity and mass density. The extrapolated distributions are evaluated through a regularization procedure relying on a finite difference estimation of the local stress tensor, as introduced in \cite{latt2008straight}.

@@

% drag signal
%
The incoming turbulent flow exerts fluctuating mechanical efforts onto the squared obstacle.
The \textit{drag} is defined as the resulting force in the streamwise $x$-direction and expresses formally as 
\begin{equation}
\label{eq:drag_definition}
f_d(t) = \int_{\mathcal{S}} \boldsymbol{\tau}_{x \beta}(\mathbf{x},t) ~ \mathrm{d}{\mathcal{S}}_\beta(\mathbf{x})  
\end{equation}
where $\mathcal{S}$ is the surface of the obstacle and $\boldsymbol{\tau}$ denotes the stress tensor (see Appendix). 
Here, the viscous stress makes a negligible contribution to the drag, which therefore results mostly from pressure forces.
% which are closely related to the distribution of velocity gradients in the vicinity of the obstacle (in the nearly-incompressible limit). 
%
Since the pressure on the top and bottom sides of the square applies in the normal direction, they do not contribute to the drag. 
As a consequence, the drag can be expressed as the pressure difference 
\begin{equation}
\label{eq:drag_approx}
f_d(t) = p_{fb}(t) - p_{base}(t)
\end{equation}
between the pressure integrated over the forebody, $p_{fb}(t)$, and the base of the obstacle, $p_{base}(t)$.

% Definition of turnover time
%
The typical timescale (turnover time) of drag fluctuations may be estimated from dimensional analysis as
\begin{equation}
\label{eq:turnover_time}
\tau_0 = \frac{R}{U}
\end{equation}
where $R$ is the size of the square and $U$ the averaged velocity. 
The viscosity is here discarded since viscous stress is negligible as compared to pressure forces.
%
% typical fluctuation of the drag
%
Fig.~\ref{fig:typical_vorticity} displays the typical evolution of the vorticity field around the obstacle over a few turnover times. It is observed that, typically, some vorticity is generated near the front part of the obstacle, rapidly detach from it and is advected downstream by the flow.
Therefore, the \textit{base pressure} varies much less than the \textit{forebody pressure} and typical drag fluctuations  result mostly from the variation of the pressure at the front of the obstacle.
In section~\ref{sec:direct_sampling}, we shall highlight that the situation leading to  {extreme} fluctuations of the drag is different.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.7\linewidth]{typical_drag_signal/typical_drag_signal.png}
  \caption{Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow. The time is normalized by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} $\tau_0=R/U$.}
  \label{fig:typical_drag_signal}
\end{figure}

@@

% Description du signal de trainee typique
Fig.~\ref{fig:typical_drag_signal} illustrates the time evolution of the drag acting on the square obstacle.
The signal appears unpredictable in details and exhibits repeated bursts of high amplitude that deviate significantly from the averaged value.
It is therefore natural to model the drag as a scalar random process $f_d(t)$.
\begin{figure}
  \centering
  \includegraphics[width=0.7\linewidth]{typical_drag_signal/typical_drag_signal.png}
  \caption{Temporal evolution of the drag acting on the square under the action of the incident turbulent flow. The time is normalized by the turnover time related the mean-flow velocity and the size of the obstacle: $\tau_0=R/U$.}
  \label{fig:typical_drag_signal}
\end{figure}

Drag fluctuations have been sampled along a long simulation of duration $T_{tot} = 4\times 10^6~\tau_0$, which will be referred to as the \textit{control run} in the following.
The corresponding estimate of the \ac{pdf} of drag fluctuations is shown in figure~\ref{fig:pdf_drag_a}.
It deviates from a normal law and shows an exponential tail for large positive fluctuations, \textit{i.e.}  ${\cal P}(f_d) = \lambda^{-1}e^{-\lambda f_d}$.
%
Fig.~\ref{fig:pdf_drag_a} also displays the \ac{pdf} of drag fluctuations acting on a control surface equivalent to the periphery of the obstacle but in the absence of the obstacle. 
%
In that case, the \ac{pdf} is quasi-symmetric and does not display exponential tails, indicating that the asymmetry of the \ac{pdf} and the development of positive exponential tail are closely related to the no-slip condition on the obstacle.
\begin{figure}
  \centering
    \subfloat[\ac{pdf} of (zero-mean) drag fluctuations]
    {\label{fig:pdf_drag_a}
    	\includegraphics[width=.45\linewidth]{./PDF_drag/PDF_drag.png}}
    \subfloat[Autocorrelation of drag fluctuations]
    {\label{fig:pdf_drag_b}
    	\includegraphics[width=.45\linewidth]{./autocorr_drag/autocorr_drag.png}}  
      \caption{\textbf{(a)} \ac{pdf} of (zero-mean) drag fluctuations $\tilde f_d \equiv f_d - \bar{f_d}$ where $\bar{f_d}$ denotes the time-averaged value. The drag is evaluated both in the presence (red) and in the absence (blue) of the obstacle. %Note that the amplitudes have not been normalized.
  \textbf{(b)} Autocorrelation function of the drag defined as $C(\tau) = \overline{ \tilde f_d(t+\tau)\tilde f_d(t)} ~/~ \overline{{\tilde f_d}^2}$. The correlation time $\tau_c$ is given by $C(\tau_c)=0$.
  %$\tau_c$ may therefore be considered as the correlation time of the drag signal.
}
  \label{fig:pdf_drag}
\end{figure}

% drag statistics
% PDF and correlation time
%
Drag fluctuations have been recorder during a simulation of duration $T_{tot} = 4 \times 10^6~\tau_0$. This very long simulation will be referred to as the \textit{control run} in the remainder. It has been made possible by the relative simplicity of the investigated two-dimensional flow and the computational efficiency of the Lattice Boltzmann method. 
%The related execution time is about a few weeks (wall-clock time).
The \ac{pdf} of the drag fluctuations is displayed in Fig.~\ref{fig:pdf_drag_a}.
It deviates from a normal law and exhibits an exponential tail for large positive values, \textit{i.e.}  ${\mathbb{P}}(f_d) \propto e^{-\lambda f_d}$.
%
Fig.~\ref{fig:pdf_drag_a} also shows the \ac{pdf} of drag fluctuations acting on a control surface corresponding the periphery of the obstacle but in the absence of the obstacle. 
%
In that latter case, the \ac{pdf} is quasi-symmetric and does not display exponential tails. This suggests that the asymmetry of the \ac{pdf} and the positive exponential tail are closely related to the no-slip condition on the obstacle.
%
%
Lastly, the autocorrelation function of the drag $C(\tau)$ is shown in Fig.~\ref{fig:pdf_drag_b}. It is found that drag fluctuations are correlated over a time interval $\tau_c \simeq 4\tau_0$ for which $C(\tau_c)=0$. Therefore, the drag looses its memory  over a time scale corresponding to the sweeping of a few eddies past the obstacle.
%
This observation will have important consequences for the application of rare-event algorithms, as will be discussed in section~\ref{sec:rare_events_algorithms}.
%
In the following, $\tau_c$ will be referred to as the \textit{correlation time} of the drag. The ratio $\tau_0 / \tau_c$ may be viewed as a {Strouhal number}; $St=0.25$ is consistent with commonly observed values for flows past blunt structures.

@@

% return time given from direct sampling
%
At first, the phenomenology of extreme fluctuations of the drag is investigated by means of \textit{direct sampling} of the control run. This brute-force approach is here used as opposed to optimized approaches involving rare-events algorithms discussed in section~\ref{sec:rare_events_algorithms}. The direct approach will provide a trustworthy baseline for the validation of rare-events algorithms.

How rare is a fluctuation $f_d \geq a$ can be quantified by its so-called \textit{return time} $r(a)$. This latter is defined as the average time between two consecutive occurrences of events of amplitudes $f_d \geq a$\TL{, as illustrated in figure~\ref{fig:illustrate_return_time}.}
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{illustrate_return_time/illustrate_return_time}
  \caption{\label{fig:illustrate_return_time} \TL{The \textit{return time} $r(a)$ of a fluctuation $f_d \geq a$ is defined as the average waiting time between the occurence of the event: $r(a) = \mathbb{E}[\tau]$. Note that we assume $r(a) >> \tau_0$, and therefore consider that such events are points in time.}}
\end{figure}
Extreme drag fluctuations are \textit{rare events} in the sense that the return time is much larger than the correlation time: $r(a) \gg \tau_c$.
In that case, the occurrence of such events follows a Poisson process \cite{lestang_computing_2018} and

@@

\begin{figure}
	\centering
	\includegraphics[width=.7\linewidth]{timeseries_extremes/timeseries_extremes.png}
	\caption{\label{fig:timeseries_extremes} Ensemble average of drag signals centered around extreme fluctuations occurring  at $t=t^{\star}$. The blue line shows the mean profile whereas the shaded area indicates variations (around the mean profile) within one standard deviation. Extreme drag events exhibit a typical lifetime of one correlation time $\tau_c$. The profile is slightly skewed indicating that the step up is slower than the return to typical values. \TL{ $\tilde f_d$ rather than $f_d$ in figure.}}
\end{figure}

>>>>

\begin{abstract}
\TL{  This study evaluates the relevance of rare-event sampling type approaches to make accessible the numerical simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
  The main idea is here to replace a long simulation by a set of much shorter simulations, running in parallel, which corresponding dynamics are replicated or pruned in order to sample large-amplitude events more frequently. 
%
  Such techniques have been shown to be efficient for a wide range of problems in statistical physics, computer science, biochemistry, enabling the simulation of events far from reach from brute force sampling approaches.
  Application to fluid-structure interaction problems remains mainly open. 
%
  The drag experienced by a squared obstacle placed in a turbulent flow (in two dimensions) is taken as a representative case study to investigate the performance of two major sampling algorithms, namely the \ac{ams} and the \ac{gktl} algorithms.
  Practical evidence is given that the fast sweeping time of fluid structures has a drastic influence on the efficiency of the sampling with both algorithms.
  While is is shown that the \ac{ams} re-sampling does not lead to rare trajectories with significant run-time savings, we illustrate that the \ac{gktl} algorithm allows for the simulation of extreme fluctuations of the time-averaged drag, as well as accurate computation of the associated statistics, for instance return times.}
\end{abstract}

@@

% general comments on physical problem %
%
Turbulent flows are important in a variety of natural phenomena, industrial and civil applications with the characteristic feature to spontaneously develop intense and sporadic motions associated with extreme internal forces \cite{lesieur_book,donzis_sreenivasan_2010,Yeung}.
``Extreme'' refers here to fluctuations that can deviate from the mean value by ${\cal{O}}(10)$ standard deviations.
In engineering, the nature of such extreme dynamical events and their statistics are of crucial interest to predict excessive mechanical efforts.
Indeed, such anomalous constraints can threaten the structural integrity of embedded structures.
%
From the viewpoint of (chaotic) dynamical systems, turbulence in fluids is linked to non-linearity and strong departure from statistical equilibrium \cite{KRAICHNAN}.
The use of perturbative methods in identifying resonant interactions among degrees of freedom responsible for extreme fluctuations is not successful.
Therefore, simulation offers a practical approach to gain physical insight into these events, quantify their intensity and estimate their frequency of occurrence.
Nevertheless, this requires very long simulations since these events are also very rare.
Our motivation is to evaluate the practicality and efficiency of rare-event sampling techniques to alleviate this prohibitive computational cost. 
%

% general def on rare-event sampling %
%
\TL{Rare-event sampling refer to a large body of methods that aim at sampling preferentially regions of phase space that correspond to rare events, that would otherwise be accessed with a very low probability through a brute-force sampling.}
%

%
% next two paragraphs may be too technical -- Find a way to summarize them in a few sentences accessible to a broad audience in fluid mechanics? Notion of action will be vague for many readers
%
% this paragraph would certainly need to be reshaped!
%
In fluid turbulence, rare-event sampling has been approached mainly from simplified dynamics such as the one-dimensional Burgers' equation with a stochastic forcing \cite{bec_burgers_2007}. In that case, dynamics may be accounted for through an action that can be identified as analogous to a Gibbs measure for a system at equilibrium. This latter can be sampled by using a Markov chain Monte-Carlo  algorithm \cite{duben_monte_2008,mesterhazy2011anomalous,mesterhazy2013lattice} that provides a framework for rare-event sampling.
%
An alternative approach is based on \emph{instantons} \cite{grafke2015instanton}. For stochastically driven systems in the limit of weak noise, it can be proved that the probability of extreme events is dominated by the minima of the action \cite{freidlin2013random} -- the trajectories that achieve such minima being referred to as instantons \cite{gurarie_instantons_1996}. Therefore, suitable numerical schemes can  be used to minimise the action and evaluate instantons as well as the related probabilities of rare events \cite{chernykh_large_2001,grafke_instanton_2013,grafke2015relevance,grigorio_instantons_2017}.
%
An obvious drawback of the aforementioned approaches is their limitation to stochastically driven dynamics which, in addition, must be simple enough so that an action may be derived and sampled, or minimised. 
%
%
%Whilst Burgers' turbulence provides an interesting model framework for the investigation of the properties of small scale turbulence, it is of little relevance to more realistic situations, such as industrial or environmental flows.
%
In this paper, a more general approach is investigated.
\TL{Je ne comprends pas cette phrase}\EL{It relies on sampling algorithms connected to the simulation of the actual dynamics itself, \textit{i.e.} applying directly to the trajectories in the dynamic space.}
% dedicated to dynamical systems for which an action cannot be easily derived or minimized,
%    
%
Even though such ideas date back to the early 1950's, they have received ever-growing interest over the last 20 years with successful applications in various domains such as chemistry \cite{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \cite{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \cite{louvin2017}, nonlinear dynamical systems \cite{tailleur_probing_2007} or communication networks simulation \cite{villen-altamirano_restart:_1994}.
%
% already mentioned in the beginning? This sentence is misleading.
%
%However, a large body of algorithms are limited to systems at equilibrium, for which the stationary measure can be sampled by means of Monte Carlo methods. 
%
An original contribution of the present study is certainly the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom. 
%Let us remark that most of previous successful applications involve dynamical systems with low-dimensional attractors. \EL{is it TRUE? \#11}
% 
Two different algorithms suitable for out-of-equilibrium dynamics are considered and compared. Namely, the Adaptive Multilevel Splitting algorithm and the Giardina-Kurchan-Tailleur-Lecomte algorithm.
%

@@

Over the last decade, the main theoretical framework for the study of rare events (in statistical physics) has been the theory of large deviations \cite{touchette_large_2009}.
Alongside, numerical methods have been developed to sample rare events \cite{moral_feynman-kac_2004}. 
Among them, the \ac{gktl} algorithm \cite{giardina_direct_2006} is particularly suited for (chaotic) dynamical systems \cite{giardina_simulating_2011,Laffargue_2013}.
%
\EL{Recently, \ac{gktl} has allowed to successfully perform the numerical simulation of extreme heat waves in a simplified modelling of the atmosphere \cite{ragone_computation_2018}. 
  This achievement represents a significant leap in the applicability of rare-event sampling to complex dynamical systems. Along the same line, rare-event sampling algorithms are here pushed aside traditional applications by considering fluid-structure interaction in a turbulent flow. }
\TL{Super!}

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]{illustr_ecoulement/illustr_ecoulement}
\caption{Our case study is a grid-generated turbulent flow impinging onto a fixed squared obstacle (of size $R$) located at the centre of a channel in two dimensions. The flow is artificially damped near the end of the channel. In the developed flow, turbulent eddies have typically the size of the square, which results in strong fluctuations of mechanical efforts acting on the square. The vorticity is displayed with an arbitrary colour map from blue (negative values) to red (positive values).}
\label{fig:illustr_ecoulement}
\end{figure}

% introduce the flow 
%
The drag exerted by a grid-generated turbulent flow onto a fixed squared obstacle is considered as a representative case study (see Fig.~\ref{fig:illustr_ecoulement}). 
% why this flow?
%
Although real-world applications would eventually imply three-dimensional dynamics, a simplified two-dimensional setting has been chosen here to reduce the computational cost and allow for a systematic study. 
% 
We believe that this system embeds the characteristic features that makes the application of rare-event algorithms both relevant and challenging for fluid-structure-interaction problems. 
%Namely, spatio-temporal chaos and the emergence of large-amplitude mechanical efforts on the structure.
%
% rapid description of the flow
Turbulent eddies generated in the near-wake of the grid are carried downstream.
They interact with each other and grow in size as expected for two-dimensional turbulent dynamics. 
The dimension of the grid is such that the size of the eddies that hit the square is comparable to its side, resulting in strong fluctuations of the drag acting on the square. 
%
%Finally, the obstacle does not deform or move either. 
%
%Through this simplified setting, our motivation is primarily to evaluate the operability of sampling techniques to capture extreme events with a significant run-time savings.  
%

@@

% geometry
%
The simulated flow develops in a long plane channel of dimension $513 \times 129$ mesh points. The squared obstacle has size $R=16$ (in mesh unit) and is located at the centre of the channel. The spacing and bar height of the entrance grid are both equal to $R/2$ (see Fig.~\ref{fig:illustr_ecoulement}). 
% boundary conditions
%
No-slip boundary conditions are enforced on top and bottom walls of the channel and on the surface of the obstacle by using an halfway bounce-back procedure \cite{lbm_book}.  
%
Upstream the grid, a constant parabolic velocity profile and a constant mass density (equal to unity) are imposed as inlet condition. 
The centerline velocity is $0.05$ in lattice unit, \textit{i.e.} normalised by $\Delta x/\Delta t$ with $\Delta x$ and $\Delta t$ referring to the lattice resolution and the time-step respectively. The distributions are imposed at equilibrium (see Appendix). 
In the bulk, the viscosity is adjusted to generate grid turbulence with Reynolds number $\mathrm{Re_{grid}}=1200$. The reference Mach number is equal to $0.06$ in agreement with the assumption of weak compressibility of the LB method. 
Near the end of the channel, the flow is progressively damped within a \textit{sponge layer} where the viscosity is artificially enhanced. Finally, the outlet boundary condition relies on a second-order extrapolation of the velocity and mass density. The extrapolated distributions are evaluated through a regularisation procedure relying on a finite difference estimation of the local stress tensor, as introduced in \cite{latt2008straight}.

@@

% drag signal
%
The incoming turbulent flow exerts fluctuating mechanical efforts onto the squared obstacle.
The \textit{drag} is defined as the resulting force in the streamwise $x$-direction, formally 
\begin{equation}
\label{eq:drag_definition}
f_d(t) = \int_{\mathcal{S}} \boldsymbol{\tau}_{x \beta}(\mathbf{x},t) ~ \mathrm{d}{\mathcal{S}}_\beta(\mathbf{x}),
\end{equation}
where $\mathcal{S}$ is the surface of the obstacle and $\boldsymbol{\tau}$ denotes the stress tensor (see Appendix). 
Here, the viscous stress makes a negligible contribution to the drag, which therefore results mostly from pressure forces.
% which are closely related to the distribution of velocity gradients in the vicinity of the obstacle (in the nearly-incompressible limit). 
%
Since the pressure on the top and bottom sides of the square applies in the normal direction, they do not contribute to the drag. 
As a consequence, the drag can be expressed as the pressure difference 
\begin{equation}
\label{eq:drag_approx}
f_d(t) = p_{fb}(t) - p_{base}(t)
\end{equation}
between the pressure integrated over the forebody, $p_{fb}(t)$, and the base of the obstacle, $p_{base}(t)$.

% Definition of turnover time
%
The typical timescale (turnover time) of drag fluctuations can be estimated from dimensional analysis as
\begin{equation}
\label{eq:turnover_time}
\tau_0 = \frac{R}{U},
\end{equation}
where $R$ is the size of the square and $U$ the averaged velocity. 
\EL{The viscosity is here discarded since viscous stress is negligible as compared to pressure forces.}\TL{Redondant avec le commentaire de (1)?}
%
% typical fluctuation of the drag
%
Fig.~\ref{fig:typical_vorticity} displays the typical evolution of the vorticity field around the obstacle over a few turnover times. It is observed that, typically, some vorticity that is generated near the front part of the obstacle rapidly detaches from it and is advected downstream by the flow.
Therefore, the \textit{base pressure} varies much less than the \textit{forebody pressure} and typical drag fluctuations mostly result from pressure variation at the front of the obstacle.
In section~\ref{sec:direct_sampling}, we shall highlight that the situation leading to  {extreme} fluctuations of the drag is different.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.7\linewidth]{typical_drag_signal/typical_drag_signal.png}
  \caption{Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow. The time is normalised by the turnover time related the mean-flow velocity and the size of the obstacle, \emph{i.e.} $\tau_0=R/U$.}
  \label{fig:typical_drag_signal}
\end{figure}

@@

% Description du signal de trainee typique
Fig.~\ref{fig:typical_drag_signal} illustrates the time evolution of the drag acting on the square obstacle.
Fluctuations appear unpredictable and the signal exhibits repeated bursts of high amplitude that deviate significantly from the averaged value.
It is therefore natural to model the drag as a scalar random process $f_d(t)$.
\begin{figure}
  \centering
  \includegraphics[width=0.7\linewidth]{typical_drag_signal/typical_drag_signal.png}
  \caption{Temporal evolution of the drag acting on the square under the action of the incident turbulent flow. The time is normalised by the turnover time related the mean-flow velocity and the size of the obstacle: $\tau_0=R/U$.}
  \label{fig:typical_drag_signal}
\end{figure}

% drag statistics
% pdf
Drag fluctuations have been sampled along a long simulation of duration $T_{tot} = 4\times 10^6~\tau_0$, which will be referred to as the \textit{control run} in the following.
It has been made possible by the relative simplicity of the investigated two-dimensional flow and the computational efficiency of the Lattice Boltzmann method. 
The corresponding estimate of the \ac{pdf} of drag fluctuations is shown in figure~\ref{fig:pdf_drag_a}.
It deviates from a normal law and shows an exponential tail for large positive fluctuations, \textit{i.e.}  ${\cal P}(f_d) = \lambda^{-1}e^{-\lambda f_d}$.
%
Fig.~\ref{fig:pdf_drag_a} also displays the \ac{pdf} of drag fluctuations acting on a control surface equivalent to the periphery of the obstacle but in the absence of the obstacle. 
%
In that case, the \ac{pdf} is quasi-symmetric and does not display exponential tails, indicating that the asymmetry of the \ac{pdf} and the development of positive exponential tail are closely related to the no-slip condition on the obstacle surface.
\begin{figure}
  \centering
    \subfloat[\ac{pdf} of (zero-mean) drag fluctuations]
    {\label{fig:pdf_drag_a}
    	\includegraphics[width=.45\linewidth]{./PDF_drag/PDF_drag.png}}
    \subfloat[Autocorrelation of drag fluctuations]
    {\label{fig:pdf_drag_b}
    	\includegraphics[width=.45\linewidth]{./autocorr_drag/autocorr_drag.png}}  
      \caption{\textbf{(a)} \ac{pdf} of (zero-mean) drag fluctuations $\tilde f_d \equiv f_d - \bar{f_d}$ where $\bar{f_d}$ denotes the time-averaged value. The drag is evaluated both in the presence (red) and in the absence (blue) of the obstacle. %Note that the amplitudes have not been normalized.
  \textbf{(b)} Autocorrelation function of the drag defined as $C(\tau) = \overline{ \tilde f_d(t+\tau)\tilde f_d(t)} ~/~ \overline{{\tilde f_d}^2}$. The correlation time $\tau_c$ is given by $C(\tau_c)=0$.
  %$\tau_c$ may therefore be considered as the correlation time of the drag signal.
}
  \label{fig:pdf_drag}
\end{figure}

% drag statistics
% correlation time
Lastly, the autocorrelation function of the drag $C(\tau)$ is shown in Fig.~\ref{fig:pdf_drag_b}. It is found that drag fluctuations are correlated over a time interval $\tau_c \simeq 4\tau_0$ for which $C(\tau_c)=0$. Therefore, the drag looses its memory  over a time scale corresponding to the sweeping of a few eddies past the obstacle.
%
This observation will have important consequences for the application of rare-event algorithms, as will be discussed in section~\ref{sec:rare_events_algorithms}.
%
In the following, $\tau_c$ will be referred to as the \textit{correlation time} of the drag. The ratio $\tau_0 / \tau_c$ may be viewed as a {Strouhal number}; $St=0.25$ is consistent with commonly observed values for flows past blunt structures.

@@

% return time given from direct sampling
%
The phenomenology of extreme fluctuations of the drag is first investigated through brute-force sampling, using the control run.
Brute-force sampling is here used as opposed to approaches involving rare-events algorithms discussed in section~\ref{sec:rare_events_algorithms}.
The direct approach will provide a trustworthy baseline for the validation of rare-events algorithms. 
How rare is a fluctuation $f_d \geq a$ can be quantified by its \textit{return time} $r(a)$. This latter is defined as the average time between two consecutive occurrences of fluctuations with amplitude $f_d \geq a$\TL{, as illustrated in figure~\ref{fig:illustrate_return_time}.}
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{illustrate_return_time/illustrate_return_time}
  \caption{\label{fig:illustrate_return_time} \TL{The \textit{return time} $r(a)$ of a fluctuation $f_d \geq a$ is defined as the average waiting time between the occurrence of the event: $r(a) = \mathbb{E}[\tau]$. Note that we assume $r(a) >> \tau_0$, and therefore consider that such events are points in time.}}
\end{figure}
Extreme drag fluctuations are \textit{rare events} in the sense that the return time is much larger than the correlation time: $r(a) \gg \tau_c$.
In that case, the occurrence of such events follows a Poisson process \cite{lestang_computing_2018} and

@@

\begin{figure}
	\centering
	\includegraphics[width=.7\linewidth]{timeseries_extremes/timeseries_extremes.png}
	\caption{\label{fig:timeseries_extremes} Ensemble average of drag signals centred around extreme fluctuations occurring  at $t=t^{\star}$. The blue line shows the mean profile whereas the shaded area indicates variations (around the mean profile) within one standard deviation. Extreme drag events exhibit a typical lifetime of one correlation time $\tau_c$. The profile is slightly skewed indicating that the step up is slower than the return to typical values. \TL{ $\tilde f_d$ rather than $f_d$ in figure.}}
\end{figure}
